Variants of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus

### 2016-04-01 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Bjordal proof fix
changed labels improves compiling

\begin{axiom}
\label{A2}
A property necessarily implied by a positive property is positive:
$$
\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
$$
\end{axiom}
\begin{theorem}
\label{T1}
Positive properties are possibly exemplified:
$$
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
$$
\end{theorem}
\begin{proof} \hfill

@@

\begin{axiom}
\label{A3}
The property of being god-like is positive:
$$
P(G)
$$
\end{axiom}

\begin{corollary}
\label{C1}
Possibly, a God-like being exists:
$$
\pos \ex x. G(x)
$$
\end{corollary}
\begin{proof} The proof of this corollary is identical to the proof of the same corollary shown in Section \ref{sec:scott}.
% \begin{prooftree}
% \AXC{Axiom \ref{A3}} \dashedLine
% \UIC{$P(G)$}
% \AXC{Theorem \ref{T1} for G} \dashedLine
% \UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
% \BIC{$\pos \ex x. G(x)$}
% \end{prooftree}
\end{proof}

@@

@@

@@

\begin{scriptsize}
\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\neg P(\rho)$} 
\AXC{Axiom 1} \dashedLine
\UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
\UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$}\RightLabel{$\imp_E$}
\BIC{$P(\lambda x. \neg \rho(x))$}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$}\RightLabel{$\imp_E$}
\BIC{$\neg \rho(\gamma)$}
\AXC{$ $} \RightLabel{3}
\UIC{$\rho(\gamma)$} \RightLabel{$\imp_E$}
\BIC{$\bot$}\RightLabel{$\imp_I^2$}
\UIC{$\neg\neg P(\rho)$}\RightLabel{$\neg\neg_E$}
\UIC{$P(\rho)$}
\AXC{Axiom \ref{A4}} \dashedLine
\UIC{$\all \varphi.[P(\varphi) \imp \nec\; P(\varphi)]$}\RightLabel{$\forall_E$}
\UIC{$P(\rho) \imp \nec\; P(\rho)$}\RightLabel{$\imp_E$}
\BIC{$\nec P(\rho)$}\RightLabel{$\imp_I^3$}
\UIC{$[\subproof_7] \qquad \rho(\gamma) \imp \nec\; P(\rho)$}
\end{prooftree}
\end{scriptsize}

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{5}
\UIC{$\nec P(\rho)$} \RightLabel{$\nec_E$}
\UIC{$\omega: P(\rho)$}
\AXC{$ $} \RightLabel{4}
\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{D1}}
\UIC{$\omega: \forall \varphi.(P(\varphi) \imp \varphi(y))$}\RightLabel{$\forall_E$}
\UIC{$\omega: P(\rho) \imp \rho(y)$}\RightLabel{$\imp_E$}
\BIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^4$}
\UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
\UIC{$\omega: \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\nec_I$}
\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^5$}
\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
\AXC{$ $} \RightLabel{6}
\UIC{$\rho(\gamma)$} \RightLabel{}
\AXC{$\subproof_7$} \dashedLine
\UIC{$\rho(\gamma)\imp\Box P(\rho)$}\RightLabel{$\imp_E$}
\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
\UIC{$\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
\UIC{$\forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$}  
\BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{D2}}
\UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$\subproof_8$} \dashedLine
\UIC{$\ess{G}{\gamma}$}
\AXC{Axiom \ref{A5}} \dashedLine
\UIC{$P(E)$}\dottedLine \RightLabel{D\ref{D3}}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(E) \imp E(\gamma)$}\RightLabel{$\imp_E$}
\BIC{$E(\gamma)$}\dottedLine \RightLabel{D\ref{D3}}
\UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
\UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{small}
\end{proof}

@@

\setcounter{theorem}{2}
\begin{theorem}
\label{T3}
Necessarily, there exists a God-like being:
$$
\nec \ex x. G(x)
$$
\end{theorem}

@@

\begin{small}
\begin{prooftree}
\AXC{ Distribution Principle} \dashedLine
\UIC{$\nec [\ex x. G(x)\imp \nec \ex x. G(x)]\imp [\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
\AXC{ Lemma \ref{L1} } \dashedLine
\UIC{$\ex x. G(x)\imp \nec \ex x. G(x)$}\RightLabel{$\nec_I$}
\UIC{$\nec[\ex x. G(x)\imp \nec \ex x. G(x)]$}\RightLabel{$\imp_E$}
\BIC{$\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
\end{prooftree}
\end{small}
\begin{small}
\begin{prooftree}
\AXC{$[\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
\AXC{ Corollary \ref{C1} } \dashedLine
\UIC{$\pos \ex x. G(x)$}\RightLabel{$\imp_E$}
\BIC{$\pos\nec \ex x. G(x)$}
\AXC{ Brouwer's Reduction Principle } \dashedLine
\UIC{$\pos\nec \ex x. G(x) \imp \ex x. G(x)$}\RightLabel{$\imp_E$}
\BIC{$ \ex x. G(x)$} \RightLabel{$\nec_I$}
\UIC{$\nec \ex x. G(x)$} 
\end{prooftree}
\end{small}

@@

\begin{theorem}
\label{T1}
If God-likeness is a positive property, then it is possibly exemplified:
$$
P(G_B) \imp \pos \ex x.G_B(x)
$$
\end{theorem}
\begin{proof} \hfill

\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\nec \neg \ex x.G_B(x) $} \RightLabel{$\nec_E$}
\UIC{$\omega: \neg \ex x.G_B(x) $} 
\AXC{$ $} \RightLabel{$1$}
\UIC{$\omega: G_B(x)$} \RightLabel{$\ex_I$}
\UIC{$\omega: \ex x.G_B(x) $} \RightLabel{$\imp_E$}
\BIC{$\omega:  \bot $}\RightLabel{$\imp_I^1$}
\UIC{$\omega:  \neg G_B(x) $} \RightLabel{$\imp_I$}
\UIC{$\omega:  G_B(x) \imp \neg G_B(x) $} \RightLabel{$\all_I$}
\UIC{$\omega:  \all x.(G_B(x) \imp \neg G_B(x)) $}  \RightLabel{$\nec_I$}
\UIC{$ \nec \all x.(G_B(x) \imp \neg G_B(x)) $}\RightLabel{$\wedge_I$}
\UIC{$P(\neg G_B)$}
\AXC{Axiom 1} \dashedLine
\UIC{$\all \varphi. [P(\neg G_B) \imp \neg P(G_B)]$} \RightLabel{$\all_E$}
\UIC{$ P( \neg G_B(x)) \imp \neg P(G_B) $}  \RightLabel{$\imp_E$}
\BIC{$\neg P(G_B) $} 
\AXC{$ $} \RightLabel{3} 
\UIC{$P(G_B)$} 
\BIC{$\bot$} \RightLabel{$\imp_I^2$}
\UIC{$ \neg \nec \neg \ex x.G_B(x)$} \doubleLine %\RightLabel{definition of $\pos$}
\UIC{$ \pos \ex x.G_B(x)$}  \RightLabel{$\imp_I^3$}
\UIC{$ P(\rho) \imp \pos \ex x.G_B(x) $} \RightLabel{$\all_I$}
\UIC{$ \all \varphi. [P(G_B) \imp \pos \ex x.G_B(x)] $}

>>>>

\begin{axiom}
\label{N:A1}
Either a property or its negation is positive, but not both:
$$
\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
$$
\end{axiom}

\begin{axiom}
\label{N:A2}
A property necessarily implied by a positive property is positive:
$$
\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
$$
\end{axiom}
\begin{theorem}
\label{N:T1}
Positive properties are possibly exemplified:
$$
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
$$
\end{theorem}
\begin{proof} \hfill

@@

\begin{definition}
\label{N:D1}
A \emph{God-like} being possesses all positive properties:
$$
G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
$$
\end{definition}

\begin{axiom}
\label{N:A3}
The property of being god-like is positive:
$$
P(G)
$$
\end{axiom}

\begin{corollary}
\label{N:C1}
Possibly, a God-like being exists:
$$
\pos \ex x. G(x)
$$
\end{corollary}
\begin{proof} The proof of this corollary is identical to the proof of the same corollary shown in Section \ref{sec:scott}.
% \begin{prooftree}
% \AXC{Axiom \ref{A3}} \dashedLine
% \UIC{$P(G)$}
% \AXC{Theorem \ref{T1} for G} \dashedLine
% \UIC{$ P(G) \imp \pos \ex x.G(x) $} \RightLabel{$\imp_E$}
% \BIC{$\pos \ex x. G(x)$}
% \end{prooftree}
\end{proof}

\begin{axiom}
\label{N:A4}
Positive properties are necessarily positive:
$$
\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
$$
\end{axiom}

\begin{definition}
\label{N:D2}
An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties:
$$
\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))
$$
\end{definition}

@@

\begin{definition}
\label{N:D3}
\emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
$$
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex y.\varphi(y)]
$$
\end{definition}

\begin{axiom}
\label{N:A5}
Necessary existence is a positive property:
$$
P(E)
$$
\end{axiom}

@@

\begin{lemma}
\label{N:L1}
If there is a God-like being, then there is a God-like being necessarily:
$$
\ex z. G(z) \imp \nec \ex x. G(x)
$$
\end{lemma}

@@

\begin{scriptsize}
\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\neg P(\rho)$} 
\AXC{Axiom 1} \dashedLine
\UIC{$\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]$} \RightLabel{$\all_E$}
\UIC{$P(\lambda x. \neg \rho(x)) \biimp \neg P(\rho)$} \RightLabel{$\wedge_E$}
\UIC{$\neg P(\rho) \imp P(\lambda x. \neg \rho(x))$}\RightLabel{$\imp_E$}
\BIC{$P(\lambda x. \neg \rho(x))$}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{N:D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(\lambda x. \neg \rho(x)) \imp \neg \rho(\gamma)$}\RightLabel{$\imp_E$}
\BIC{$\neg \rho(\gamma)$}
\AXC{$ $} \RightLabel{3}
\UIC{$\rho(\gamma)$} \RightLabel{$\imp_E$}
\BIC{$\bot$}\RightLabel{$\imp_I^2$}
\UIC{$\neg\neg P(\rho)$}\RightLabel{$\neg\neg_E$}
\UIC{$P(\rho)$}
\AXC{Axiom \ref{N:A4}} \dashedLine
\UIC{$\all \varphi.[P(\varphi) \imp \nec\; P(\varphi)]$}\RightLabel{$\forall_E$}
\UIC{$P(\rho) \imp \nec\; P(\rho)$}\RightLabel{$\imp_E$}
\BIC{$\nec P(\rho)$}\RightLabel{$\imp_I^3$}
\UIC{$[\subproof_7] \qquad \rho(\gamma) \imp \nec\; P(\rho)$}
\end{prooftree}
\end{scriptsize}

\begin{small}
\begin{prooftree}
\AXC{$ $} \RightLabel{5}
\UIC{$\nec P(\rho)$} \RightLabel{$\nec_E$}
\UIC{$\omega: P(\rho)$}
\AXC{$ $} \RightLabel{4}
\UIC{$\omega: G(y)$}  \dottedLine \RightLabel{D\ref{N:D1}}
\UIC{$\omega: \forall \varphi.(P(\varphi) \imp \varphi(y))$}\RightLabel{$\forall_E$}
\UIC{$\omega: P(\rho) \imp \rho(y)$}\RightLabel{$\imp_E$}
\BIC{$\omega: \rho(y)$} \RightLabel{$\imp_I^4$}
\UIC{$\omega: G(y) \imp \rho(y)$}\RightLabel{$\forall_I$}
\UIC{$\omega: \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\nec_I$}
\UIC{$\nec \forall y.(G(y) \imp \rho(y))$} \RightLabel{$\imp_I^5$}
\UIC{$\nec P(\rho)\imp \nec \forall y .(G(y)\imp \rho(y))$} \RightLabel{$\imp_E$}
\AXC{$ $} \RightLabel{6}
\UIC{$\rho(\gamma)$} \RightLabel{}
\AXC{$\subproof_7$} \dashedLine
\UIC{$\rho(\gamma)\imp\Box P(\rho)$}\RightLabel{$\imp_E$}
\BIC{$\nec P(\rho)$} \RightLabel{$\imp_E$}
\BIC{$\nec \forall y .(G(y)\imp \rho(y))$}\RightLabel{$\imp_I^6$}
\UIC{$\rho(\gamma) \imp \nec \forall y.(G(y) \imp \rho(y))$}\RightLabel{$\forall_I$}
\UIC{$\forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$}  
\BIC{$G(\gamma) \wedge \forall \psi.(\psi(\gamma) \imp \nec \forall y.(G(y) \imp \psi(y)))$}\dottedLine \RightLabel{D\ref{N:D2}}
\UIC{$[\subproof_8] \qquad \ess{G}{\gamma}$}
\end{prooftree}
\end{small}

\begin{small}
\begin{prooftree}
\AXC{$\subproof_8$} \dashedLine
\UIC{$\ess{G}{\gamma}$}
\AXC{Axiom \ref{A5}} \dashedLine
\UIC{$P(E)$}\dottedLine \RightLabel{D\ref{N:D3}}
\AXC{$\subproof_6$} \dashedLine
\UIC{$G(\gamma)$}\dottedLine \RightLabel{D\ref{N:D1}}
\UIC{$\forall \varphi.(P(\varphi) \imp \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(E) \imp E(\gamma)$}\RightLabel{$\imp_E$}
\BIC{$E(\gamma)$}\dottedLine \RightLabel{D\ref{N:D3}}
\UIC{$ \all \varphi.[\ess{\varphi}{\gamma} \imp \nec \ex x.\varphi(x)] $}\RightLabel{$\all_E$}
\UIC{$ \ess{G}{\gamma} \imp \nec \ex x. G(x) $}\RightLabel{$\imp_E$}
\BIC{$\nec \ex x. G(x)$} \RightLabel{$\imp_I^1$}
\UIC{$\ex z. G(z) \imp \nec \ex x. G(x)$}
\end{prooftree}
\end{small}
\end{proof}

@@

\setcounter{theorem}{2}
\begin{theorem}
\label{N:T3}
Necessarily, there exists a God-like being:
$$
\nec \ex x. G(x)
$$
\end{theorem}

@@

\begin{small}
\begin{prooftree}
\AXC{ Distribution Principle} \dashedLine
\UIC{$\nec [\ex x. G(x)\imp \nec \ex x. G(x)]\imp [\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
\AXC{ Lemma \ref{N:L1} } \dashedLine
\UIC{$\ex x. G(x)\imp \nec \ex x. G(x)$}\RightLabel{$\nec_I$}
\UIC{$\nec[\ex x. G(x)\imp \nec \ex x. G(x)]$}\RightLabel{$\imp_E$}
\BIC{$\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)$}
\end{prooftree}
\end{small}
\begin{small}
\begin{prooftree}
\AXC{$[\pos \ex x. G(x)\imp \pos\nec \ex x. G(x)]$}
\AXC{ Corollary \ref{C1} } \dashedLine
\UIC{$\pos \ex x. G(x)$}\RightLabel{$\imp_E$}
\BIC{$\pos\nec \ex x. G(x)$}
\AXC{ Brouwer's Reduction Principle } \dashedLine
\UIC{$\pos\nec \ex x. G(x) \imp \ex x. G(x)$}\RightLabel{$\imp_E$}
\BIC{$ \ex x. G(x)$} \RightLabel{$\nec_I$}
\UIC{$\nec \ex x. G(x)$} 
\end{prooftree}
\end{small}

@@

\begin{theorem}
\label{B:T1}
If God-likeness is a positive property, then it is possibly exemplified:
$$
P(G_B) \imp \pos \ex x.G_B(x)
$$
\end{theorem}
\begin{proof} \hfill

\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\nec \neg \ex x.G_B(x) $} \RightLabel{$\nec_E$}
\UIC{$\omega: \neg \ex x.G_B(x) $} 
\AXC{$ $} \RightLabel{$1$}
\UIC{$\omega: G_B(x)$} \RightLabel{$\ex_I$}
\UIC{$\omega: \ex x.G_B(x) $} \RightLabel{$\imp_E$}
\BIC{$\omega:  \bot $}\RightLabel{$\imp_I^1$}
\UIC{$\omega:  \neg G_B(x) $} \RightLabel{$\imp_I$}
\UIC{$\omega:  G_B(x) \imp \neg G_B(x) $} \RightLabel{$\all_I$}
\UIC{$\omega:  \all x.(G_B(x) \imp \neg G_B(x)) $}  \RightLabel{$\nec_I$}
\UIC{$ \nec \all x.(G_B(x) \imp \neg G_B(x)) $}\RightLabel{$\wedge_I$}
\UIC{$P(\neg G_B)$}
\AXC{Axiom 1} \dashedLine
\UIC{$\all \varphi. [P(\neg G_B) \imp \neg P(G_B)]$} \RightLabel{$\all_E$}
\UIC{$ P( \neg G_B(x)) \imp \neg P(G_B) $}  \RightLabel{$\imp_E$}
\BIC{$\neg P(G_B) $} 
\AXC{$ $} \RightLabel{3} 
\UIC{$P(G_B)$} 
\BIC{$\bot$} \RightLabel{$\imp_I^2$}
\UIC{$ \neg \nec \neg \ex x.G_B(x)$} \doubleLine %\RightLabel{definition of $\pos$}
\UIC{$ \pos \ex x.G_B(x)$}  \RightLabel{$\imp_I^3$}
\UIC{$ P(G_B) \imp \pos \ex x.G_B(x) $}
